---
title: "class Vis 1 onclass"
author: "yingyu cui"
output: "github_document"
---
StackOverflow has a lot of questions and answers.

graph components : geoms?  facets scales statistics 
what is ridge graph 

load data
```{r}
library(tidyverse)
library(ggridges)
```
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```
 make first plot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```
make sure to subl=mit everything 
filter(is.na()) function 

aes(x=, y=, color=, )
geom_point (alpha= tranparent= .3, size = .8)
geom_smooth(se=FALSE)   smooth curve 
 this is diifferent with below 
aes(x=, y=,)
geom_point (aes= (color=), alpha= tranparent= .3, size = .8)
the smooth curve will become from 3 to 1 

use faceting function 
```{r}
facet_grid(. ~ name)
facet_grid(name ~ .)
```
size = prcp  the size of each point is determined by the value of prcp  


weather_df |> 
filter |> 
  mutate() |> 
  ggplot geom_smooth(method = "lm", se = FALSE) no curve line just straight one
  
geom_hex function
ggplot color function 
why color = blue turns every point red?  color = "blue" turns the color blue, with out "" not 

univariable 
ggplot(color = name/ fill = name)
geom_histogram(position = "dodge")
have a try 
maybe the facet_grid fuction is better could have a try 

geom_density(alpha= ) density plot 

geom_boxplot()

violin plots 
geom_violin()

reorder function 

ridge plot 
geom_density_ridges() in library(ggridges)  need to install the library first
bimodal or monomoal 

filter function 

save and embed plots
ggsave("filename.XXX", the poltname above, width = ,height = )


embed picture 
fig.width fig.asp
how to change the size of the picture 

