---
title: "class_vis1_afterclass"
author: "Yingyu Cui"
date: "2024-10-07"
output: github_document
---

```{r set up}
library(tidyverse)
library(ggridges)
```
```{r download and create weather data frame}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
view(weather_df)
```

Begin with scatterplot
```{r}
weather_df |>
ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```
```{r scatterplot with pipe function}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

another way to plot
```{r another way to plot}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()
```

name function
```{r name function}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```
smooth curve and point transparent
```{r smooth curve and point transparent}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```
where to place the name function and facet grid
note the place of "coler=name" which determine whether the color of lines are defferent by name.
facet_grid(.~ name)---vertical graphs by name, facet_grid(name ~ .)---horizontal graphs by name
```{r where to place the name function and facet grid}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
geom_plot size function
```{r geom_plot size function}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

assessment:
```{r assessment}
weather_df |> 
  filter(name == "CentralPark_NY") |>
  mutate(tmin = tmin * 9/5 + 32, tmax = tmax * 9/5 + 32) |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_smooth(method = "lm", se = FALSE)
```

geom_hex() function 
visualization when number of data is too big and scatterplot is not clear when using transparency function
geom_bin2d(), or geom_density2d() can be used as well
```{r geom_hex() function}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

difference between these: one is regulate the coler of points, the other is to create a new varable
```{r difference between}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```

univariable
geom_histogram()function
attention: fill = name not color = name
```{r univariable histogram}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()
```
more aesthetics: dodge make sorting not overlapping 
```{r more aesthetics}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```
The position = "dodge" places the bars for each group side-by-side, but this gets sort of hard to understand. I often prefer density plots in place of histograms.
```{r density plots}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```
adjust is like binwidth, but for density plots. 
 You could facet by name as above but would have to ask if that makes comparisons easier or harder.

add geom_rug() for better show raw data in addition to density plot
show the distribution situation of raw data
```{r geom_rug()}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue") +
  geom_rug()
```

Histograms and densities are one way of investigating univariate distributions; boxplots are another.
```{r boxplots}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_boxplot()
```

Violin plots are a combination of boxplots and density plots.
stat_summary() could add some information on the plot
```{r violin plots}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

 if you have lots of categories in which the shape of the distribution matters---- ridge plot 
```{r ridge plot}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

use ggsave() by explicitly creating the figure and exporting
(using relative paths!).
ggsave("ggp_weather.pdf", ggp_weather(or path = relative path), width = 8, height = 5)

### **`out.width` 和 `fig.width` 的区别**：
- **`fig.width` 和 `fig.height`**：
  - 这些参数设置的是图形**在生成时**的实际尺寸，影响的是图形的分辨率。
  - 设置较大的 `fig.width` 和 `fig.height` 可以生成更高分辨率的图形。

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
can be inserted in the beginning of the Rmd file.


What makes embedding figures difficult at first is that things like the font and point size in the figures generated by R are constant – that is, they don’t scale with the overall size of the figure. As a result, text in a figure with width 12 will look smaller than text in a figure with width 6 after both have been embedded in a document.
