---
title: "class_vis1_afterclass"
author: "Yingyu Cui"
date: "2024-10-08"
output: html_document
---

---
title: "class_vis1_afterclass"
author: "Yingyu Cui"
date: "2024-10-07"
output: github_document
---

```{r set up}
library(tidyverse)
library(ggridges)
```
```{r download and create weather data frame}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
view(weather_df)
```

Begin with scatterplot
```{r scatterplot}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r scatterplot with pipe function}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

another way to plot
```{r another way to plot}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()
```

name function
```{r name function}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```
smooth curve and point transparent
```{r smooth curve and point transparent}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```
where to place the name function and facet grid
note the place of "coler=name" which determine whether the color of lines are defferent by name.
facet_grid(.~ name)---vertical graphs by name, facet_grid(name ~ .)---horizontal graphs by name
```{r where to place the name function and facet grid}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```
geom_plot size function
```{r geom_plot size function}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

assessment:
```{r assessment}
weather_df |> 
  filter(name == "CentralPark_NY") |>
  mutate(tmin = tmin * 9/5 + 32, tmax = tmax * 9/5 + 32) |>
  ggplot(aes(x = tmin, y = tmax)) +
  geom_smooth(method = "lm", se = FALSE)
```

geom_hex() function 
visualization when number of data is too big and scatterplot is not clear when using transparency function
geom_bin2d(), or geom_density2d() can be used as well
```{r geom_hex() function}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

difference between
```{r difference between}
ggplot(weather_df) + geom_point(aes(x = tmax, y = tmin), color = "blue")
```