---
title: "EDA_afterclass"
author: "Yingyu Cui"
date: "2024-10-08"
output: github_document
---

# Exploratory analysis using data summaries
Conduct exploratory analyses using dplyr verbs (group_by and summarize), along with numeric data summaries

```{r set up}
library(tidyverse)
```

```{r load data}
weather_df_EDA = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) |>
  select(name, id, everything())
```

initial numeric exploration
```{r histogram}
weather_df |> 
  ggplot(aes(x = prcp)) + 
  geom_histogram()
```
some data visualization should be helpful before following explorations, if clearly skewed, we should pay attention.



group_by function
could follow by summarize function or mutate function
```{r group_by}
weather_df_EDA |>
  group_by(name, month)
```
it is sometimes necessary to remove groups using ungroup()



counting 
 count the number of observations in each month
```{r count}
weather_df_EDA |>
  group_by(month) |>
  summarize(n_obs = n())
```
在 R 中，summarize() 是 dplyr 包中的一个函数，用于对数据进行汇总计算。它通常与 group_by() 一起使用，可以在分组数据的基础上计算汇总统计量，如平均值、总和、计数、最小值、最大值等
the result is a dataframe that includes the grouping variable(s) and the desired summary.
```{r anothor way to count}
weather_df |>
  count(month, name = "n_obs")
```
n_days = n_distinct(date) 
n_distinct() 是 dplyr 包中的一个函数，用于计算一个向量或列中的唯一值的数量。
换句话说，它会去重，并计算不重复的值的数量


2X2 table
using group_by and summarize to create a 2X2 table
```{r 2X2 table}
weather_df |> 
  drop_na(tmax) |> 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""
  )) |> 
  filter(name != "Molokai_HI") |> 
  group_by(name, cold) |> 
  summarize(count = n())
```
re-organize into a more standard (non-tidy) 2x2 table using pivot_wider, or you could use janitor::tabyl() to create a 2x2 table
{r reorganise}
janitor::tabyl(name, cold)

janitor has a lot of little functions like this that turn out to be useful, so when you have some time you might read through all the things you can do. 

Just do a logistic regression adjusting for confounders 
sometimes this table is not that useful, we can directly do a logistic regression to adjust for confounders




General summerization
summarize() using functions like mean(), median(), var(), sd(), mad(), IQR(), min(), and max()
indicate the variable to which they apply and include any additional arguments as necessary
If you want to summarize multiple columns using the same summary, the across function is helpful.
  summarize(across(tmin:prcp, mean))

The fact that summarize() produces a dataframe 
is important (and consistent with other functions in the tidyverse). You can incorporate grouping and summarizing within broader analysis pipelines.

The results of group_by() and summarize() are generally tidy, but presenting reader-friendly results for this kind of exploratory analysis often benefits from some un-tidying. 

using 
pivot_wider(
    names_from = name,
    values_from = mean_tmax) |> 
  knitr::kable(digits = 1)



Grouped function
group_by() and mutat() could be used 
weather_df |>
  group_by(name) |>
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax)
    


Window function
Window functions, in contrast, take n inputs and return n outputs, and the outputs depend on all the inputs. 
There are several categories of window functions; you’re most likely to need ranking functions and offsets

{r window function}
weather_df |>
  group_by(name, month) |>
  mutate(temp_ranking = min_rank(tmax)) |> 
  filter(min_rank(tmax) < 2) |> 
  
  desc() 是一个函数，用于对数值进行降序排序
    filter(min_rank(desc(tmax)) < 4)
    
Offsets, especially lags, are used to compare an observation to it’s previous value.
which is used for the variability of data with time changing
or to identify the largest one-day increase
{r offset}
weather_df |>
  group_by(name) |>
  mutate(temp_change = tmax - lag(tmax))



