---
title: "vis2_afterclass"
author: "Yingyu Cui"
date: "2024-10-08"
output: html_document
---

```{r set up}
library(tidyverse)
library(patchwork)
```

load weather data
```{r load data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
weather_df
```

scatterplot of tmax and tmin
```{r scatterplot}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

use of labels to show some details for others
the use of caption and color
```{r labels}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package"
  )
```

control the specification and location of tick marks on the X or Y axis.
scale_x_continuous() and scale_y_continuous() or change continuous to discrete
breaks and labels
```{r axis}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15"))
```

other useages:
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ºC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
For instance, scale_y_sqrt() can be added to a ggplot object to transform the Y scale, and xlim() can be used to control the plot limits in the X axis.



control the color
  scale_color_hue(h = c(100, 300))
I encourage you to use the viridis package instead. There are several options, but the default color scheme works nicely!

viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
 The viridis::scale_fill_viridis() function is appropriate for the fill aesthetic used in histograms, density plots, and elsewhere.
We used discrete = TRUE because the color aesthetic is mapped to a discrete variable. In other cases (for example, when color mapped to prcp) you can omit this argument to get a continuous color gradient



Themes

modify non-data elements of a plot – they don’t change mappings or how data are render, but control things like background color and location of the the legend. Using themes can help with general plot appearance.
By default this is on the right of the graphic, but I like to shift it to the bottom to ensure the graphic takes up the available left-to-right space.

ggp_temp_plot + 
  theme(legend.position = "bottom")
legend.position = "none" will remove the legend. 

This is helpful when multiple plots use the same color scheme or when the legend is obnoxious for some other reason.

for the background for plots:
 By default this is theme_gray(); 
 here’s theme_bw()---a white background with black lines and text---and theme_minimal()---a white background with gray lines and text.
 and here’s theme_classic()---a white background with black lines and text and no gridlines.
 the Excel 2003 theme from ggthemes:
 ggthemes::theme_excel()
 Attention:
 Don’t use the Excel 2003 theme (the first two are fine, and ggthemes has other very nice themes as well).
 
 The ordering of theme_bw() and theme() matters theme_bw() will override the theme() settings, so it should come last.
 

```{r figure preferrences}

library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
these could be copy pasted in every beginning of Rmd file.




overlay data on a complete data figure
```{r overlay}
central_park_df = 
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(data = molokai_df, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park_df) 
```
one way to do this is to create a “summary” dataframe and use that when adding a new geom to a ggplot based on the full data


Patchwork
a solution is to create each of the panels you want separately and combine panels using tools in the patchwork package
```{r patchwork}
tmax_tmin_p = 
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```
The package is already very helpful but is under active development – some features may change or be added over time – so check the package webstie periodically to stay up-to-date. 





data manipulation
in some cases, it’s easier to control behavior through data manipulation than it is through the plot code.

for the order of categorical or factor variables, 
Categorical variables will be ordered alphabetically; factors will follow the specified order level that underlies the variable labels. You can change the order level of a factor variable to your specified preference using forcats::fct_relevel 
or according to the value of another variable using forcats::fct_reorder.

{r reorder catergorical variables to specified preferrence}
mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA")))

{r reorder catergorical variables to the value of another variable}
mutate(name = forcats::fct_reorder(name, tmax))


reorder the data when tidying the data or manipulating the data
{r reorder data}
 mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) |>
  arrange(id, visit)
  
  recode function and factor function are used to reorder the data.

 mutate(outcome = forcats::fct_reorder(outcome, pn_day, median))
 pn_day's median value is used to reorder the data.
 
 



 



